### 2015-08-05 whoop.tex
Minor tweaks to save a line here and there

Initially, the valuation of shared variables is arbitrary, no mutexes are held (i.e.\ each mutex maps to $\bot$), and the list of threads is empty.
%
At any point of execution, a new thread may be added to the list of threads.  This involves selecting a thread template $T$ and an id $i \in \ids$ that has not been previously used during program execution, setting the point of execution for the new thread to be the first statement of $\mainforthread{T}$, and choosing an arbitrary valuation for the private variables $\variables{T}$.
%
We consider a standard interleaving model of concurrency: at any execution point, a thread may execute its current statement, unless that statement has the form $\mutexlock{m}$ and mutex $m$ is held by some thread.  Executing a statement causes the state of the thread, and the shared state, to be updated in a standard manner.  For example, if a thread following template $T$ executes $s = e$, where $s \in \sharedvars$ and $e$ is an expression over $\variables{T}$, the shared variable valuation is updated so that $s$ has the value determined by evaluating $e$ in the context of the thread's private variable valuation.  Because our interest is in data race analysis for race-free programming, we are not concerned with relaxed memory behavior: race-free programs exhibit only sequentially consistent behaviors.

@@

\noindent\textbf{Sequentialization }
%
The pseudocode of Fig.~\ref{fig:sequentialization} shows the sequential program that we analyze in order to prove race-freedom for a pair of thread templates $T$ and $U$. Assuming that $T$ and $U$ have been instrumented using distinct symbols $i$ and $j$, yielding $\instrument{T}{i}$ and $\instrument{U}{j}$, the sequential program operates as follows. First, the read, write, and current locksets for $\instrument{T}{i}$ and $\instrument{U}{j}$ are initialized to be empty, and for each shared variable $s$, the locksets $\lockset{s}{i}$ and $\lockset{s}{j}$ are initialized to the full set of mutexes $\mutexes$.  The main procedures of the instrumented thread templates, $\mainforthread{\instrument{T}{i}}$ and $\mainforthread{\instrument{U}{j}}$, are then executed in turn (the order does not matter due to renaming). Finally, an assertion checks for consistent use of mutexes: if $s$ is written during execution of $\instrument{T}{i}$ and accessed during execution of $\instrument{U}{j}$, or vice-versa, then the locksets $\lockset{s}{i}$ and $\lockset{s}{j}$ must contain at least one common mutex.

@@

\end{tabular}
\caption{The sequential program to be analyzed in order to prove race-freedom for a pair of thread templates}
\label{fig:sequentialization}
%\vspace{-2mm}
\end{figure}

>>>>

Initially, the valuation of shared variables is arbitrary, no mutexes are held (i.e.\ each mutex maps to $\bot$), and the list of threads is empty.
%
At any point of execution, a new thread may be added to the list of threads.  This involves selecting a thread template $T$ and an id $i \in \ids$ that has not been previously used during program execution, setting the point of execution for the new thread to be the first statement of $\mainforthread{T}$, and choosing an arbitrary valuation for the private variables $\variables{T}$.
%
We consider a standard interleaving model of concurrency: at any execution point, a thread may execute its current statement, unless that statement has the form $\mutexlock{m}$ and mutex $m$ is held by some thread.  Executing a statement updates the thread-private and shared state in a standard manner.  For example, if a thread following template $T$ executes $s = e$, where $s \in \sharedvars$ and $e$ is an expression over $\variables{T}$, the shared variable valuation is updated so that $s$ has the value determined by evaluating $e$ in the context of the thread's private variable valuation.  Because our interest is in data race analysis for race-free programming, we are not concerned with relaxed memory behavior: race-free programs exhibit only sequentially consistent behaviors.

@@

\noindent\textbf{Sequentialization }
%
Fig.~\ref{fig:sequentialization} shows the sequential program that we analyze to prove race-freedom for a pair of thread templates $T$ and $U$. Assuming that $T$ and $U$ have been instrumented using distinct symbols $i$ and $j$, yielding $\instrument{T}{i}$ and $\instrument{U}{j}$, the sequential program operates as follows. First, the read, write, and current locksets for $\instrument{T}{i}$ and $\instrument{U}{j}$ are initialized to be empty, and for each shared variable $s$, the locksets $\lockset{s}{i}$ and $\lockset{s}{j}$ are initialized to the full set of mutexes $\mutexes$.  The main procedures of the instrumented thread templates, $\mainforthread{\instrument{T}{i}}$ and $\mainforthread{\instrument{U}{j}}$, are then executed in turn (the order does not matter due to renaming). Finally, an assertion checks for consistent use of mutexes: if $s$ is written during execution of $\instrument{T}{i}$ and accessed during execution of $\instrument{U}{j}$, or vice-versa, then the locksets $\lockset{s}{i}$ and $\lockset{s}{j}$ must contain at least one common mutex.

@@

\end{tabular}
\caption{The sequential program to be analyzed to prove race-freedom for a pair of thread templates}
\label{fig:sequentialization}
%\vspace{-2mm}
\end{figure}
